---
title: "DATA605 HW1"
author: "Misha Kollontai"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Problem set 1

#### (1) Show that $A^{T}A \neq AA^{T}$ in general. (Proof and demonstration.)

Let's look at the example of a 3x3 matrix A

$$\mathbf{A} = \left[\begin{array}
{rrr}
n_{11} & n_{12} & n_{13} \\
n_{21} & n_{22} & n_{23} \\
n_{31} & n_{32} & n_{33}
\end{array}\right]
$$
This would mean that:

$$\mathbf{A^{T}} = \left[\begin{array}
{rrr}
n_{11} & n_{21} & n_{31} \\
n_{12} & n_{22} & n_{32} \\
n_{13} & n_{23} & n_{33}
\end{array}\right]$$

Let's take a look at what the first column will look like for $A^{T}A$:

$$\mathbf{A^{T}A[,1]} = \left[\begin{array}
{r}
n_{11} * n_{11} + n_{21} * n_{21} + n_{31} * n_{31} \\
n_{12} * n_{11} + n_{22} * n_{21} + n_{32} * n_{31} \\
n_{13} * n_{11} + n_{23} * n_{21} + n_{33} * n_{31}
\end{array}\right]$$

Let's take a look at what the first column will look like for $AA^{T}$:

$$\mathbf{AA^{T}[,1]} = \left[\begin{array}
{r}
n_{11} * n_{11} + n_{12} * n_{12} + n_{13} * n_{13} \\
n_{21} * n_{11} + n_{22} * n_{12} + n_{23} * n_{13} \\
n_{31} * n_{11} + n_{32} * n_{12} + n_{33} * n_{13}
\end{array}\right]$$

#### As we can see from the examples above, generally $A^{T}A \neq AA^{T}$ There are cases where the coefficients may match up, but usually this will not be true. 

To test this, let's look at a random 4x4 matrix A:

```{r test ATA_AAT}
#Define a 4x4 matrix A using random values
A <- c(1,2,-5,6,-1,-7,3,5,4,3,9,13,-5,3,-12,7)
dim(A) <- c(4,4)

#Transpose this matrix and assign to AT
AT <- t(A)

#Calculate the two multiplied matrices
ATA <- AT %*% A

AAT <- A %*% AT

ATA == AAT
```

**As we can see, none of the entries in the two matrices match, confirming our proof.**

---

#### (2) For a special type of square matrix $A$, we get $A^{T}A = AA^{T}$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

This could be true in the case that $A = A^{T}$. The identity matrix is one such example of a diagonal matrix.

---

## 2. Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

```{r Prob_2}

findLU <- function(x){
  n <- nrow(x)
  L <- diag(n)
  
  for (i in c(2:n)){
    for (j in c(1:(i-1))){
      Elim <- (x[i,j]/x[j,j])
      L[i,j] <- Elim
      x[i,] <- x[i,]-Elim*x[j,]
    }
  }
  print("The matrix L =")
  print(L)
  return(x)
}

A <- c(1,2,3,1,1,1,2,0,1)
dim(A) <- c(3,3)
print("The matrix A =")
print(A)
U <- findLU(A)
print("The matrix U =")
U

```