---
title: "DATA605 HW3"
author: "Misha Kollontai"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Problem set 1

#### (1) What is the rank of the matrix $A$?

$$\mathbf{A} = \left[\begin{array}
{rrrr}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3
\end{array}\right]$$

``` {r Prob1_1}
require(pracma)

A <- c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3)
dim(A) <- c(4,4)
A

rref(A)
```

The RREF of matrix A above shows us that it has a rank of 4 (4 pivotS). To confirm, use R to calculate the rank of A:

```{r}
Rank(A)
```

#### (2) Given an $mxn$ matrix where $m>n$, what can be the maximum rank? The minimum rank assuming the matrix is non-zero?

The maximum rank can be $n$ and the minimum rank can be 1.

#### (3) What is the rank of matrix $B$?

$$\mathbf{B} = \left[\begin{array}
{rrr}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2 
\end{array}\right]$$

``` {r Prob1_3}

B <- c(1,3,2,2,6,4,1,3,2)
dim(B) <- c(3,3)
B

rref(B)
```

The RREF of matrix B above shows us that it has a rank of 1 (1 pivot). To confirm, use R to calculate the rank of B:

```{r}
Rank(B)
```

## 2. Problem Set 2

#### Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

```{r}
A <- c(1,0,0,2,4,0,3,5,6)
dim(A) <- c(3,3)
```

$$\mathbf{A} = \left[\begin{array}
{rrr}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6 
\end{array}\right]$$

To get the eigenvalues:

$$det(A -\lambda I) =  0$$

$$\mathbf{(A - \lambda I)} = \left[\begin{array}
{rrr}
1-\lambda & 2 & 3 \\
0 & 4-\lambda & 5 \\
0 & 0 & 6-\lambda
\end{array}\right]$$

$$det(A-\lambda I) = (1-\lambda)*((4-\lambda)*(6-\lambda)-0) - 2*(0) + 3*(0)$$
$$det(A-\lambda I) = (1-\lambda)*(24-10\lambda+\lambda^{2})$$

$$det(A-\lambda I) = 24 - 10\lambda+\lambda^{2}-24\lambda+10\lambda^{2}-\lambda^{3}$$
$$-\lambda{3}+11\lambda^{2}-34\lambda+24 = 0$$
```{r}
require(RConics)
det <- c(-1,11,-34,24)
eigVal <- cubic(det)

#The eigenvalues of matrix A are:
eigVal
```

**First Eigenvalue: $\lambda_{1} = 1$**

$$(A-\lambda_{1} I)\nu=0$$
$$\mathbf{(A - \lambda_{1} I)} = \left[\begin{array}
{rrr}
0 & 2 & 3 \\
0 & 3 & 5 \\
0 & 0 & 5
\end{array}\right]$$

$$0(\nu_{1}) + 2(\nu_{2})+3(\nu_{3}) =0$$
$$2(\nu_{2})=-3(\nu_{3})$$
$$\nu_{2}=-3/2(\nu_{3})$$
$$0(\nu_{1}) + 0(\nu_{2}) + 5(\nu_{3})=0$$
$$\nu_{3} = 0$$
$$\nu_{2} = 0$$
$$\nu_{1} - anything$$

**Therefore:**

$$\nu = (1,0,0)$$
**Second Eigenvalue: $\lambda_{2} = 6$**

$$(A-\lambda_{2} I)\nu=0$$
$$\mathbf{(A - \lambda_{2} I)} = \left[\begin{array}
{rrr}
-5 & 2 & 3 \\
0 & -2 & 5 \\
0 & 0 & 0
\end{array}\right]$$

$$-5(\nu_{1}) + 2(\nu_{2})+3(\nu_{3}) =0$$
$$0(\nu_{1})-2(\nu_{2})+5(\nu_{3})=0$$
$$2\nu_{2}=5\nu_{3}$$
$$\nu_{2}=2.5(\nu_{3})$$
$$-5(\nu_{1})+8(\nu_{3})=0$$
$$8\nu_{3}=5\nu_{1}$$
$$\nu_{3} = 5/8(\nu_{1})$$
$$\nu_{2} = 2.5(\nu_{3})$$

**Assuming $\nu_{1}=1$:**

$$\nu = (1,1.5625,0.625)$$
**Converting to unit vector:**

```{r}
unitize <- function(x){
  x / sqrt(sum(x^2))
}
vec1 <- c(1,1.5625,0.625)

unit1 <- unitize(vec1)
unit1
```

$$\nu = (0.5108, 0.7982, 0.3193)$$

**Third Eigenvalue: $\lambda_{3} = 4$**

$$(A-\lambda_{3} I)\nu=0$$
$$\mathbf{(A - \lambda_{3} I)} = \left[\begin{array}
{rrr}
-3 & 2 & 3 \\
0 & 0 & 5 \\
0 & 0 & 2
\end{array}\right]$$

$$-3(\nu_{1}) + 2(\nu_{2})+3(\nu_{3}) =0$$
$$5(\nu_{3})=0$$
$$\nu_{3} = 0$$
$$3(\nu_{1}) = 2(\nu_{2})$$
$$\nu_{1} = 2/3(\nu_{2})$$
**Assuming $\nu_{2} = 1$:**

$$\nu = (0.6667,1,0)$$
**Converting to unit vector:**

```{r}
vec3 <- c(0.6667,1,0)

unit3 <- unitize(vec3)
unit3
```

#### The three eigenvalue/eigenvector pairs for matrix A are:
(1) $\lambda_{1} = 1$
$$\nu = (1,0,0)$$
(2) $\lambda_{2} = 6$
$$\nu = (0.5108, 0.7982, 0.3193)$$
(3) $\lambda_{3} = 4$
$$\nu = (0.5547,0.8320,0)$$

