---
title: "DATA605 HW1"
author: "Misha Kollontai"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Problem set 1

You can think of vectors representing many dimensions of related information. For
instance, Netflix might store all the ratings a user gives to movies in a vector. This is
clearly a vector of very large dimensions (in the millions) and very sparse as the user might
have rated only a few movies. Similarly, Amazon might store the items purchased by a user
in a vector, with each slot or dimension representing a unique product and the value of the
slot, the number of such items the user bought. One task that is frequently done in these
settings is to find similarities between users. And, we can use dot-product between vectors
to do just that. As you know, the dot-product is proportional to the length of two vectors
and to the angle between them. In fact, the dot-product between two vectors, normalized
by their lengths is called as the cosine distance and is frequently used in recommendation
engines.

#### (1) Calculate the dot product $u.v$ where $u=[0.5;0.5]$ and $v=[3;-4]$

```{r Prob 1_1}
u <- c(0.5,0.5)
v <- c(3,-4)
z <- u %*% v
z
```

The dot product $u.v$ is `r z`.

---

#### (2) What are the lengths of $u$ and $v$? Please note that the mathematical notion of the length of a vector is not the same as a computer science definition.

```{r prob 1_2}
length_u <- sqrt(0.5^2 + 0.5^2)
length_v <- sqrt(3^2 + (-4)^2)
```

The length of $u =$ `r length_u` and the length of $v=$ `r length_v`.

---

#### (3) What is the linear combination: $3u − 2v$?

```{r Prob 1_3}
Prob1_3 <- 3*u - 2*v
```

The linear combination $3u - 2v$ is [`r Prob1_3`]

---

#### (4) What is the angle between $u$ and $v$?

```{r Prob 1_4}
require(matlib)
cos_a <- z / (length_u*length_v)
alpha <- angle(u,v,FALSE)
alpha_2 <- acos(cos_a)
if (alpha == alpha_2){
  print("Both calculations yielded same result")
}
deg_alpha <- alpha * 180 / pi
```

The angle $\alpha$ between $u$ and $v$ is equal to `r alpha` rad or `r deg_alpha` deg.

---

## 2. Problem set 2

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. You don’t have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve to solve this system or use matrix inverses. The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.

```{r Prob 2}
require(pracma)

UppTri <- function(x,y){
  Aug_base <- cbind(x,y)
  levels <- c(0,0,0)
  Aug <- Aug_base
  
  for(i in c(1:3)){
    if (Aug[i,1] != 0){levels[i] <- 1}
    else {if(Aug[i,2] != 0){levels[i] <- 2}
       else{levels[i] <- 3}
    }
  }

## Start by sorting the rows in order of present variables
   if(levels[1] > levels[2]){
     if(levels[2] < levels[3]){
       Aug[1,] <- Aug_base[2,]
       Aug[2,] <- Aug_base[1,]
       Aug[3,] <- Aug_base[3,]
     } else{
       if(levels[3] <= levels[2]){
          Aug[1,] <- Aug_base[3,]
          Aug[2,] <- Aug_base[2,]
          Aug[3,] <- Aug_base[1,]
       } else{
          Aug[1,] <- Aug_base[2,]
          Aug[2,] <- Aug_base[3,]
          Aug[3,] <- Aug_base[1,]
       }
     }
   } else{
     if(levels[3] > levels[1]){
        Aug[1,] <- Aug_base[1,]
        Aug[2,] <- Aug_base[3,]
        Aug[3,] <- Aug_base[2,]
     } else{
       if(levels[3] < levels[2]){
          Aug[1,] <- Aug_base[3,]
          Aug[2,] <- Aug_base[1,]
          Aug[3,] <- Aug_base[2,]
       }
     }
   }
  
  #Set the first term to 1
  if(Aug[1,1] != 0){
    Aug[1,] <- Aug[1,]/Aug[1,1]
  }
  
  #Ensure that the other two rows have zero values in the first column
  Aug[2,] <- Aug[2,] - Aug[1,]*Aug[2,1]/Aug[1,1]
  Aug[3,] <- Aug[3,] - Aug[1,]*Aug[3,1]/Aug[1,1]
  
  #Set the first value in the second row equal to 1
  if(Aug[2,2] != 0){
    Aug[2,] <- Aug[2,]/Aug[2,2]
    #Zero the value in the third row below the 1 of the second
    Aug[3,] <- Aug[3,] - Aug[2,]*Aug[3,2]/Aug[2,2]
    #If possible, do the same to the first row
    if(Aug[1,2] != 0){
      Aug[1,] <- Aug[1,] - Aug[2,]*Aug[1,2]/Aug[2,2]
    }
  }
  
  if(Aug[3,3] != 0){
    Aug[3,] <- Aug[3,]/Aug[3,3]
    if(Aug[1,3] != 0){
      Aug[1,] <- Aug[1,] - Aug[3,]*Aug[1,3]/Aug[3,3]
    }
    if(Aug[2,3] != 0){
      Aug[2,] <- Aug[2,] - Aug[3,]*Aug[2,3]/Aug[3,3]
    }
  }
  fail <- FALSE
  for(i in c(2,3)){
    if(Aug[i,1] == 0 & Aug[i,2] == 0 & Aug[i,3] ==0 && Aug[i,4] != 0){
      fail <- TRUE
    }
  }
  
  if (fail){
    print("There is no solution")
  }else{
    print("The solution to the system of equations using my function is:")
    print(Aug[,4])
  }
  return(Aug[,4])
}

## check that the function finds the correct answer

A <- c(1,2,-1,1,-1,-2,3,5,4)
dim(A) <- c(3,3)
b <- c(1,2,6)

print(A)
mySolution <- UppTri(A,b)
RREF <- rref(cbind(A,b))
RREF2 <- solve(A,b)

print("The solutions calculated by built-in R functions are:")
RREF[,4]
RREF2

print("The solutions match!")
```